Local Spectral Clustering of Density Upper Level Sets

### 2019-10-04 arxiv/local_spectral_density_clustering_arxiv.tex
added relevant examples tuff

\subsection{PPR on a neighborhood graph} 
We now describe the clustering
algorithm that will be our focus for the rest of the paper. Let $\Xbf = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $\Xbf$, an unweighted, undirected graph with
vertices $V=\Xbf$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
x_j} \leq r$, where $\norm{\cdot}$ is the $\ell_2$ norm. We denote by $\Abf \in
\Reals^{n \times n}$ the adjacency matrix, with entries $\Abf_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $\Dbf$ the diagonal degree
matrix, with $\Dbf_{uu} = \sum_{v \in V} \Abf_{uv}$, and by $\Ibf{}$ the $n
\times n$ identity matrix.

Next, we define the PPR vector $\pbf = \pbf(v,\alpha;G_{n,r})$, based on
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, to be 
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
\pbf = \alpha \ebf{v} + (1 - \alpha) \pbf \Wbf,
\end{equation}
where $\Wbf = (\Ibf{} + \Dbf^{-1}\Abf)/2$ is the lazy random walk matrix over
$G_{n,r}$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in the
$v$-th position and 0 elsewhere). For a level $\beta > 0$ and a target volume
$\vol_0 > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf = (p_u)_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta := \set{u \in V: \frac{p_u}{\Dbf_{uu}} > \frac{\beta}{\vol_{0}}}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{11})$}\footnote{The choice of a specific range such as  
\smash{$(\frac{1}{40}, \frac{1}{11})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

\begin{algorithm}
\caption{PPR on a neighborhood graph}
\label{alg: ppr}	
{\bfseries Input:} data $\Xbf=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
parameter $\alpha \in [0,1]$, seed $v \in \Xbf$, target stationary volume
$\vol_0 > 0$. \\     
{\bfseries Output:} cluster $\Cest \subseteq V$.
\begin{algorithmic}[1]
  \STATE Form the neighborhood graph $G_{n,r}$.
  \STATE Compute the PPR vector $p=\pbf(v, \alpha; G_{n,r})$ as in \eqref{eqn: 
    ppr_vector}. 
  \STATE For \smash{$\beta \in (\frac{1}{40}, \frac{1}{11})$} compute sweep cuts 
  $S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
  \STATE Return the cluster \smash{$\Cest = S_{\beta^*}$}, where  
  $$
  \beta^* = \argmin_{\beta \in (\frac{1}{40}, \frac{1}{11})}~ \Phi(S_{\beta}; G_{n,r}).
  $$
\end{algorithmic}
\end{algorithm}

@@

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. \citet{shi2009} and \citet{schiebinger2015} examine the consistency of  
spectral algorithms in recovering the latent labels in certain
nonparametric mixture models. Their results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

More broadly, density clustering and level set estimation is a well-studied
problem. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

@@

  %%% AJG 5/20: Should I turn this into two conditions?
\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
  \begin{enumerate}
  \item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
  with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
  \diam < \infty$
  \item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
  $\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$ and
  \item for some $L
  \geq 1$,   
  \begin{equation*}
    \frac{1}{L}\norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
  \end{equation*}
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite diameter 
  under a  measure preserving, bi-Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  We assume that neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
    2 \int_{\Csig} \Pbb(B(x,r)) f(x) dx \leq \int_{\Rd} \Pbb(B(x,r)) f(x) dx,
  \end{equation*}
  where $B(x,r)$ is the closed ball of radius $r$ at $x$.
\end{enumerate}
To motivate these conditions, we provide a brief high-level sketch
of our analysis (we return to this more formally in Section~\ref{sec: analysis}).
\citet{zhu2013} show for an arbitrary graph $G =
(V,E)$ and subset of vertices $S \subseteq V$, the PPR estimate \smash{$\Cest$}
of subset $S$ satisfies, for a constant $c>0$,
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G)
\cdot \tau_{\infty}(G[S])\bigr) \vol(S;G),
\end{equation}
where $\Phi(S;G)$ is the normalized cut of $S$ (as defined in \eqref{eqn:
  normalized_cut}), and \smash{$\tau_{\infty}(G[S])$} is  the \emph{mixing
  time} of a random walk over the induced subgraph $G[S]$ (to be defined
precisely later, in \eqref{eqn: mixing_time}).  The left-hand side in
\eqref{eqn:  graph_symmetric_set_difference_1} resembles a (degree-weighted) 
form of the symmetric set difference metric in \eqref{eqn:
  misclassification_rate}, and our main goal will be to upper bound 
  this quantity. 
  Towards this goal, as we will show in Section \ref{sec:
  analysis}, the conditions \ref{asmp: bounded_density}--\ref{asmp:
  bounded_volume} allow us to upper bound the normalized cut 
\smash{$\Phi(\Csig[\Xbf]; G_{n,r})$}, and the mixing time 
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. In more detail, assumptions  
\ref{asmp: cluster_separation}, \ref{asmp: low_noise_density} yield an upper
bound on \smash{$\cut(\Csig[\Xbf]; G_{n,r})$}, and \ref{asmp: bounded_density}
yields a lower bound on \smash{$\vol(\Csig[\Xbf]; G_{n,r})$}; together with 
\ref{asmp: bounded_volume}, this gives an upper bound on the normalized cut.  On 
the other hand, \ref{asmp: bounded_density}, \ref{asmp: embedding} preclude
bottlenecks in the induced subgraph \smash{$G_{n,r}[\Csig[\Xbf]]$}, and combined
with the upper bound on diameter in \ref{asmp: embedding}, this leads to an upper bound on
the mixing time over this subgraph.

@@

\begin{definition}[Well-conditioned density clusters]
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some
  $\theta$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
  later, we set 
  \begin{equation}
    \label{eqn: condition_number_1}
    \Phibf(\theta) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}},~  
    \taubf(\theta) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3,
  \end{equation}
  and letting \smash{$\kappa(\Cset) := \Phi_{u}(\theta) \cdot
    \tau_{u}(\theta)$}, we call $\Cset$ a 
  \emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}
\noindent The condition number 
$\kappa(\Cset)$ succinctly captures the role of the various
geometric parameters. 
We note in passing that $\Phibf(\theta)$ and $\taubf(\theta)$ are exactly the upper bounds 
on \smash{$\Phi(\Csig[\Xbf]; G_{n,r})$} and
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in our analysis
later, in Section \ref{sec: analysis}.

\subsubsection{Well-initialized algorithm} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$ (with respect to some $\theta$), we
require that the tuning parameters of Algorithm~\ref{alg: ppr} are chosen as
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9] \cdot \frac{1}{\taubf(\theta)}, \\
v \in \Csig[\Xbf]^g, \quad
\vol_0 \in [3/4,5/4] \cdot n(n-1) \int_{\Csig} \Pbb(B(x,r)) f(x) dx, 
\end{gathered}
\end{equation}
where \smash{$\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$} will be some large
(``good'') subset of $\Csig[\Xbf]$. In particular, abbreviating $\vol_{n,r}(S)
:= \vol(S; G_{n,r})$ for $S \subseteq \Xbf$, this good subset has large volume, i.e. 
$\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$.

\begin{definition}
If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
  initialization} for some well-conditioned density cluster $\Cset$, we
say the algorithm is \emph{well-initialized}. 
\end{definition}
\noindent In practice it is  not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and selects the cluster which has the smallest normalized cut. 
%it is not obvious how this scheme would affect the performance of PPR in the
%density clustering context.

\subsection{Cluster Recovery in Symmetric Set Difference}
The results of Section \ref{sec: analysis}, combined 
with \eqref{eqn: graph_symmetric_set_difference_1}, give an upper bound on the
volume of \smash{$\Cest \setminus \Csig[\Xbf]$} and \smash{$\Csig[\Xbf]
  \setminus \Cest$},  
\begin{equation}
\label{eqn: graph_symmetric_set_difference}
\vol_{n,r}(\Cest \setminus \Csig[\Xbf]) + \vol_{n,r}(\Csig[\Xbf] \setminus
\Cest) \leq c \kappa(\Cset) \vol_{n,r}(\Csig[\Xbf]). 
\end{equation}
To translate the guarantee in~\eqref{eqn: graph_symmetric_set_difference} into meaningful bounds
on the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf], \Cest)$}, we
want to preclude vertices $x \in \Xbf$ from having arbitrarily small degree, and
so we make some regularity assumptions on $\mathcal{X} := \mathrm{supp}(f)$. Let
$\nu$ denote the Lebesgue measure on $\Rd$, and $\nu_d := \nu(B)$ be the measure
of the unit ball $B = B(0,1)$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{5}
\item 
  \label{asmp: valid_region}
  \emph{Regular support:} There exists some constant $\lambda_{\min} > 0$ such 
  that $\lambda_{\min} < f(x)$ for all $x \in \mathcal{X}$. Additionally, there
  exists some $c > 0$ such that for each $x \in \partial \mathcal{X}$,
  $\nu(B(x,r) \cap \mathcal{X}) \geq c \nu_d r^d$. 
\end{enumerate}
We note that the latter condition in \ref{asmp: valid_region} will hold if the
boundary $\partial \mathcal{X}$ is sufficiently regular.  Now we present our
first main result: a bound on the symmetric set difference metric.
\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume that $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}
The proof of Theorem \ref{thm: misclassification_rate} is deferred to Appendix~\ref{sec: proof_of_misclassification_rate}.
This result shows that the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

>>>>

\subsection{PPR on a neighborhood graph} 
We now describe the clustering
algorithm that will be our focus for the rest of the paper. Let $\Xbf = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density~$f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $\Xbf$, an unweighted, undirected graph with
vertices $V=\Xbf$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
x_j} \leq r$, where $\norm{\cdot}$ is the Euclidean norm. We denote by $\Abf \in
\Reals^{n \times n}$ the adjacency matrix, with entries $\Abf_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $\Dbf$ the diagonal degree
matrix, with $\Dbf_{uu} = \sum_{v \in V} \Abf_{uv}$, and by $\Ibf{}$ the $n
\times n$ identity matrix.

Next, we define the PPR vector $\pbf_v = \pbf(v,\alpha;G_{n,r})$, based on
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, to be 
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
\pbf_v = \alpha \ebf{v} + (1 - \alpha) \pbf \Wbf,
\end{equation}
where $\Wbf = (\Ibf{} + \Dbf^{-1}\Abf)/2$ is the lazy random walk matrix over
$G_{n,r}$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in the
$v$-th position and 0 elsewhere).

\subsection{Sweep cut.}
\label{subsection: sweep_cut}

Finally, we employ the sweep cut method to obtain a local cluster from the PPR embedding $\pbf_v$. For a level $\beta > 0$ and a target stationary probability $\pi_0 > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf_v = (\pbf_v(u))_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta := \set{u \in V: \pbf_v(u) > \beta \pi_0}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{12})$}\footnote{The choice of a specific range such as  
\smash{$(\frac{1}{40}, \frac{1}{12})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, the PPR algorithm is summarized in Algorithm~\ref{alg: ppr}.

\begin{algorithm}
\caption{PPR on a neighborhood graph}
\label{alg: ppr}	
{\bfseries Input:} data $\Xbf=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
parameter $\alpha \in [0,1]$, seed $v \in \Xbf$, target stationary probability
$\pi_0 > 0$. \\     
{\bfseries Output:} cluster $\Cest \subseteq V$.
\begin{algorithmic}[1]
  \STATE Form the neighborhood graph $G_{n,r}$.
  \STATE Compute the PPR vector $\pbf_v=\pbf(v, \alpha; G_{n,r})$ as in \eqref{eqn: 
    ppr_vector}. 
  \STATE For \smash{$\beta \in (\frac{1}{40}, \frac{1}{12})$} compute sweep cuts 
  $S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
  \STATE Return the cluster \smash{$\Cest = S_{\beta^*}$}, where  
  $$
  \beta^* = \argmin_{\beta \in (\frac{1}{40}, \frac{1}{11})}~ \Phi(S_{\beta}; G_{n,r}).
  $$
\end{algorithmic}
\end{algorithm}

@@

\subsection{Related Work and Our Contributions}
In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. In the stochastic block model (SBM), arguably one of the simplest models of network formation, edges between nodes independently occur with probability based on a latent community membership. The ability of spectral clustering to perform community detection, in this context is well-understood. In particular, \citet{rohe2011} upper bound the fraction of nodes misclassified by a spectral algorithm for the high-dimensional (large number of blocks) SBM,  \citet{lei2015} extend these results to the sparse (low average degree) regime, and \citet{balakrishnan2011} analyze the misclassification rate when the block model exhibits some hierarchical structure. The framework we consider, where nodes correspond to data points sampled from an underlying density, and edges between nodes are formed based on geometric proximity, is quite different than the SBM, and therefore so is our analysis.

The study of spectral algorithms on neighborhood graphs has focused on establishing consistency results-- that is, asymptotic convergence of eigenvalues and eigenvectors of matrices to eigenvalues and eigenfunctions of limiting operators. \citet{koltchinskii2000} establish convergence of spectral projections of the adjacency matrix to a limiting integral operator, with similar results obtained using simplified proofs in \citet{rosasco10}. \citet{vonluxburg2008} studies convergence of eigenvectors of the Laplacian matrix for a neighborhood of fixed radius. \citet{belkin07} and \citet{garciatrillos18} extend these results to the regime $r \to 0$ as $n \to \infty$.

These results are of fundamental importance; however, the behavior of the spectra of these underlying integral operators is somewhat hard to grasp. Therefore, further work relating this spectra to the geometry of the distribution $\Pbb$ is of interest. In this spirit, \citet{shi2009} and \citet{schiebinger2015} examine the ability of  
spectral algorithms to recover the latent labels in certain geometrically well-conditioned nonparametric mixture models. Their results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

A substantial portion of our technical effort involves estimating graph-theoretic connectivity properties of empirical density clusters in the neighborhood graph $G_{n,r}$. In particular, we prove an upper bound on the mixing time of a random walk run only over the subset of nodes in $G_{n,r}$ which fall within the empirical density cluster $\Cset[\Xbf]$. To do so, we rely on a series of seminal works upper bounding the mixing time of \emph{geometric random walks} (see \citet{vempala2005} for a comprehensive review.)  \citet{dyer1991} initiated this study, using these random walks as a fundamental subroutine to efficiently compute volumes of high-dimensional convex bodies. These results are improved in \citep{lovasz1990,kannan97,kannan06}, who show, among other things, that the bounds on mixing time can be sharpened by avoiding so-called ``start-penalties''.(These improvements are crucial to our work, for they allow us to avoid factors of $\log n$ which would imply divergence as $n \to \infty$.) Following the lead of \citep{abbasi-yadkori2016,abbasi-yadkori2016a}, we extend these results to hold for Lipschitz deformations of convex sets. Additionally, we relate the mixing time of these (continuous-space) geometric random walks to the mixing time of random walks over (discrete) neighborhood graphs.

Finally, it is worth mentioning that density clustering and level set estimation is of course a well-studied problem. \citet{polonik1995, rigollet2009} study density clustering under the 
symmetric set difference metric, \citet{tsybakov1997, singh2009} describe
minimax optimal level-set estimators under Hausdorff loss and
\citet{hartigan1981, chaudhuri2010, balakrishnan2013} consider consistent estimation of the
cluster tree.
%, to note but a few works. 
We emphasize that our goal is not to improve on these
results, nor to offer a better algorithm for level set estimation; indeed, seen as
a density clustering algorithm, PPR has none of the optimality guarantees 
found in the aforementioned works. Instead, our motivation is to start with a 
widely-used local spectral method, PPR, and to better understand and
characterize the distinctions between those density clusters which are
well-conditioned for PPR, and those which are not.

@@

\item In Section \ref{sec: examples}, we show the implications of our theory for some simple generative models, and compare our upper bounds on the symmetric set difference metric to those achievable by classical density cluster estimators.

@@

\item
  \label{asmp: embedding}
  \emph{Lipschitz embedding:}
  There exists $g: \Reals^d \to \Reals^d$ with the following properties: 
  \begin{enumerate}
  \item we have $\Csig = g(\mathcal{K})$, for a convex set $\mathcal{K} \subseteq \Rd$
  with $\mathrm{diam}(\mathcal{K}) = \sup_{x,y \in \mathcal{K}}\norm{x - y} =:
  \diam < \infty$
  \item $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, where
  $\nabla g(x)$ is the Jacobian of $g$ evaluated at $x$ and
  \item for some $L
  \geq 1$,   
  \begin{equation*}
    \norm{g(x) - g(y)} \leq L \norm{x - y} ~
    \text{for all $x,y \in \mathcal{K}$}. 
  \end{equation*}
  \end{enumerate}
  Succinctly, we assume that $\Csig$ is the image of a convex set with finite diameter 
  under a  measure preserving, Lipschitz transformation.

\item
  \label{asmp: bounded_volume}
  \emph{Bounded volume:}
  We assume that neighborhood graph radius $0 < r \leq \sigma/2d$ is chosen such that
  \begin{equation*}
    2 \int_{\Csig} \Pbb(B(x,r)) f(x) dx \leq \int_{\Rd} \Pbb(B(x,r)) f(x) dx,
  \end{equation*}
  where $B(x,r)$ is the closed ball of radius $r$ at $x$.
\end{enumerate}

To motivate these conditions, we provide a brief high-level sketch
of our analysis (we return to this more formally in Section~\ref{sec: analysis}). 
\citet{zhu2013} show for an arbitrary graph $G =
(V,E)$ and subset of vertices $S \subseteq V$, the PPR estimate \smash{$\Cest$}
of subset $S$ satisfies, for a constant $c>0$,
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G)
\cdot \tau_{\infty}(G[S])\bigr),
\end{equation}
where $\Phi(S;G)$ is the normalized cut of $S$ (as defined in \eqref{eqn:
  normalized_cut}), and \smash{$\tau_{\infty}(G[S])$} is  the \emph{mixing
  time} of a random walk over the induced subgraph $G[S]$ (to be defined
precisely later, in \eqref{eqn: mixing_time}).  The left-hand side in
\eqref{eqn: graph_symmetric_set_difference_1} resembles a (degree-weighted) 
form of the symmetric set difference metric in \eqref{eqn:
  misclassification_rate}, and our main goal will be to upper bound 
  this quantity. 
  Towards this goal, as we will show in Section \ref{sec:
  analysis}, the conditions \ref{asmp: bounded_density}--\ref{asmp:
  bounded_volume} allow us to upper bound the normalized cut 
\smash{$\Phi(\Csig[\Xbf]; G_{n,r})$}, and the mixing time 
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$}. In more detail, assumptions  
\ref{asmp: cluster_separation}, \ref{asmp: low_noise_density} yield an upper
bound on \smash{$\cut(\Csig[\Xbf]; G_{n,r})$}, and \ref{asmp: bounded_density}
yields a lower bound on \smash{$\vol(\Csig[\Xbf]; G_{n,r})$}; together with 
\ref{asmp: bounded_volume}, this gives an upper bound on the normalized cut.  On 
the other hand, \ref{asmp: bounded_density}, \ref{asmp: embedding} preclude
bottlenecks in the induced subgraph \smash{$G_{n,r}[\Csig[\Xbf]]$}, and combined
with the upper bound on diameter in \ref{asmp: embedding}, this leads to an upper bound on
the mixing time over this subgraph.

\textcolor{red}{Mention Arias-Castro and Little here.}

@@

\begin{definition}[Well-conditioned density clusters]
  \label{def:well_conditioned_density_cluster}
  For $\lambda > 0$ and \smash{$\Cset \in \Cbb_f(\lambda)$}, let $\Cset$ satisfy 
  \ref{asmp: bounded_density}--\ref{asmp: bounded_volume} for some
  $\theta$. Then, for universal constants $c_1, c_2, c_3 > 0$ to be specified
  later, we set 
  \begin{equation}
    \label{eqn: condition_number_1}
    \Phibf(\theta) 
    := c_1 r \frac{d}{\sigma} \frac{\lambda}{\lambda_{\sigma}}
    \frac{(\lambda_{\sigma} - c_0 \frac{r^{\gamma}}{\gamma +
        1})}{\lambda_{\sigma}},~  
    \taubf(\theta) := c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2
      L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3,
  \end{equation}
  and letting \smash{$\kappa(\Cset) := \Phi_{u}(\theta) \cdot
    \tau_{u}(\theta)$}, we call $\Cset$ a 
  \emph{$\kappa$-well-conditioned density cluster}.  
\end{definition}
\noindent The condition number 
$\kappa(\Cset)$ succinctly captures the role of the various
geometric parameters. 
We note in passing that $\Phibf(\theta)$ and $\taubf(\theta)$ are exactly the upper bounds 
on \smash{$\Phi(\Csig[\Xbf]; G_{n,r})$} and
\smash{$\tau_{\infty}(G_{n,r}[\Csig[\Xbf]])$} that we derive in our analysis
later, in Section \ref{sec: analysis}.

\subsubsection{Well-initialized algorithm} As is typical in the local clustering
literature, our algorithmic results will be stated with respect to specific
ranges of each of the user-specified parameters. In particular, for a
well-conditioned density cluster $\Cset$ (with respect to some $\theta$), we require that the tuning parameters of Algorithm~\ref{alg: ppr} are chosen to fall within the ranges
\begin{equation}
\begin{gathered}
\label{eqn: initialization}
0 < r \leq \frac{\sigma}{2d}, \quad 
\alpha \in [1/10, 1/9] \cdot \frac{1}{\taubf(\theta)}, \\
v \in \Csig[\Xbf]^g, \quad
\pi_0 \in [3/4,5/4] \cdot \frac{\lambda_{\sigma}}{n\Lambda_{\sigma}\Pbb(\Csig)},
\end{gathered}
\end{equation}
where \smash{$\Csig[\Xbf]^g \subseteq \Csig[\Xbf]$} will be some large
(``good'') subset of $\Csig[\Xbf]$. In particular, abbreviating $\vol_{n,r}(S)
:= \vol(S; G_{n,r})$ for $S \subseteq \Xbf$, this good subset has large volume, i.e. 
$\vol_{n,r}(\Csig[\Xbf]^g) \geq \vol_{n,r}(\Csig[\Xbf])/2$.

\begin{definition}
If the input parameters to Algorithm \ref{alg: ppr} satisfy \eqref{eqn:
  initialization} for some well-conditioned density cluster $\Cset$, we
say the algorithm is \emph{well-initialized}. 
\end{definition}
\noindent In practice it is  not feasible to set hyperparameters based on the
underlying (unknown) density $f$. Typically, one tunes PPR over a range of
hyperparameters and selects the cluster which has the smallest normalized cut.

\subsection{Cluster Recovery in Symmetric Set Difference}
We now present our first main result: a bound on the symmetric set difference metric between the estimated cluster $\widehat{C}$ and empirical cluster $\Csig[\Xbf]$.
\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$ and let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$). If Algorithm \ref{alg: ppr} is well-initialized, there exists a universal constant $c > 0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)  \frac{\Lambda_{\sigma}}{\lambda_{\sigma}} \abs{\Csig[\Xbf]}
  \end{equation}
\end{theorem}
The proof of Theorem \ref{thm: misclassification_rate} is deferred to Appendix~\ref{sec: proof_of_misclassification_rate}.
This result shows that the symmetric set difference metric \smash{$\Delta(\Csig[\Xbf],
  \Cest)$} is upper-bounded by a quantity proportional to the difficulty of the clustering problem, as
measured by the condition number $\kappa(\Cset)$.

\subsection{Consistent Cluster Recovery}
The bound on symmetric set difference \eqref{eqn: misclassification_rate_ub} does not imply consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}. This notion of
consistency requires a uniform bound over $\pbf$: for all \smash{$\Cset'
  \in \Cbb_f(\lambda), \Cset' \neq \Cset$}, and each $u \in \Cset, w \in
\Cset'$,  
\begin{equation}
\label{eqn: ppr_gap}
\pbf_v(w) \leq \frac{\pi_0}{40} < \frac{\pi_0}{11} \leq \pbf_v(u),
\end{equation}
so that any sweep cut $S_{\beta}$ for $\beta \in [1/40,1/11]$ (i.e., any
sweep cut considered by Algorithm \ref{alg: ppr}) will fulfill both conditions
laid out in \eqref{eqn: consistent_density_cluster_recovery}. In Theorem
\ref{thm: consistent_recovery_of_density_clusters}, we show that a sufficiently
small upper bound on $\kappa(\Cset)$ ensures that as $n \to \infty$, such a gap exists with probability
1, and hence guarantees \smash{$\Cest$} will be a consistent  
estimator. We will need one additional regularity condition, to preclude arbitrarily low degree
vertices for points $x \in \Cset'[\Xbf]$. 
\begin{enumerate}[label=(A\arabic*)]
  \setcounter{enumi}{6}
\item 
  \label{asmp: C'_bounded_density}
  \emph{Bounded density in other clusters:} Letting $\sigma,\lambda_{\sigma}$ be 
  as in \ref{asmp: bounded_density}, for each $\Cset' \in \Cbb_f(\lambda)$ and
  for all $x \in \Csig'$, $\lambda_{\sigma} \leq f(x)$. 
\end{enumerate}
